Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-05-11 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Fixed simulation times of Berger-DPPC-98 in Table 1 to reported (60/30ns): http://nmrlipids.blogspot.com/2014/11/new-version-of-manuscript.html?showComment=1421231975768#c3582830273528800239
udpate CHARMM36 liteture information in table

Berger-DPPC-98~\cite{marrink98}    & DPPC &  72 & 2864 & 323 & 140 & 100 & [\citenum{bergerDPPCfiles}]$^*$ & SI \\

@@

CHARMM36~\cite{klauda10}  & DPPC &  72 & 2189 & 323 &  -  & 130 & -   & [\citenum{klauda10}]$^h$  \\

@@

MacRog~\cite{maciejewski14}  & POPC & 288 & 12600 & 310 & 100 &  80 & [\citenum{macrogFILES}]$^*$      & SI  \\

@@

MacRog}~\cite{maciejewski14} & POPC & 128 &  6400 & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$  & SI  \\
MacRog~\cite{maciejewski14}  & POPC & 288 & 14400 & 310 &  90 &  40 & [\citenum{macrogdehydFILES}]$^*$ & SI  \\

@@

\begin{table}[htb]
\centering
\caption{Simulated single component lipid bilayers with varying hydration levels. The simulation file data sets marked with $^*$ include also part of the trajectory.
$^a$ Water/lipid molar ratio
$^b$ The number of lipid molecules
$^c$ The number of water molecules
$^d$ Simulation temperature
$^e$ The total simulation time
$^f$ Time frames used in the analysis
$^g$ Reference link for the downloadable simulation files
$^h$ Reference for the full simulation details
}\label{systemsDEHYD}
\begin{tabular}{c c c c c c c c c c}
%\hline
Force field & lipid & $^a$n (w/l)   & $^b$N$_{\rm l}$   &  $^c$N$_{\rm w}$ & $^d$T (K)  & $^e$t$_{{\rm sim}}$(ns)  & $^f$t$_{{\rm anal}}$ (ns)& $^g$Files  &  $^h$Details\\
\hline
Berger-POPC-07~\cite{ollila07a}          &   POPC & 57  &128 & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & SI \\
                                        &   POPC & 7  &128 & 896   & 298  & 60 & 50 & [\citenum{bergerDEHYDfiles}]$^*$ & SI \\
Berger-DLPC-13~\cite{kanduc13}          &   DLPC & 28  &72 & 2016  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc28}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 24  &72 & 1728  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc24}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 20  &72 & 1440  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc20}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 16  &72 & 1152  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc16}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 12  &72 & 864  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc12}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 8  &72 & 576  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc8}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 4  &72 & 288  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc4}]$^*$ & [\citenum{kanduc13}] \\
CHARMM36\cite{klauda10}          & POPC   & 40 & 128 &  5120   & 303 & 150 & 100  & [\citenum{charmm36files}]$^*$   & SI \\
                                 & POPC   & 31  & 72  &  2242 & 303 & 30 & 20 & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
                               & POPC   & 15 & 72 &  1080  & 303 & 59 & 40 & [\citenum{charmm36files15wPERl}]$^*$ & SI \\
                            & POPC   & 7  & 72  &  504  & 303 & 60 & 20 & [\citenum{charmm36files7wPERl}]$^*$ & SI \\
MacRog\cite{maciejewski14}     & POPC   & 50 & 288  & 14400 & 310 & 90 & 40 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                               & POPC   & 25 & 288  & 7200 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 20 & 288  & 5760 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 15 & 288  & 4320 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 10 & 288  & 2880 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 5 & 288   & 1440 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
GAFFlipid\cite{dickson12}      & POPC   & 31& 126  & 3948  & 303 & 137 & 32 & [\citenum{GAFFlipidFILES}]$^*$ & SI  \\
                               & POPC   & 7& 126  & 896   & 303 & 130 & 40 & [\citenum{gaffDEHYDfiles}]$^*$ & SI \\
\end{tabular}
\end{table}

\begin{sidewaystable}[]
\centering
\caption{Simulated lipid bilayers containing cholesterol. The simulation file data sets marked with $^*$ include also part of the trajectory.
$^a$ The number of lipid molecules
$^b$ The number of cholesterol molecules
$^c$ Cholesterol concentration (mol\%)
$^d$ The number of water molecules
$^e$ Simulation temperature
$^f$ The total simulation time
$^g$ Time frames used in the analysis
$^h$ Reference link for the downloadable simulation files
$^i$ Reference for the full simulation details
}\label{systemsCHOL}
\begin{tabular}{c c c c c c c c c c c}
%\hline
Force field & lipid   & $^a$N$_{\rm l}$ & $^b$N$_{\rm chol}$ &$^c$C$_{\rm CHOL}$  &  $^d$N$_{\rm w}$ & $^e$T (K)  & $^f$t$_{{\rm sim}}$(ns)  & $^g$t$_{{\rm anal}}$ (ns)& $^h$Files  &  $^i$Details\\
\hline
Berger-POPC-07~\cite{ollila07a}&   POPC &128 & 0 &0\% & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & [\citenum{ferreira15}] \\
/H\"oltje-CHOL-13~\cite{holtje01,ferreira13}   &    & &  &   &   &  &  &  &  \\
                               &   POPC &120 & 8 & 6\% &7290   & 298  & 100 & 80 & [\citenum{bergerFILESpopc7chol}]$^*$ & [\citenum{ferreira13}] \\
                               &   POPC &110 & 18& 14\% & 8481  & 298  & 100 & 80 & [\citenum{bergerFILESpopc15chol}]$^*$ & [\citenum{ferreira13}]  \\
                               &   POPC &84 & 44 & 34\%  & 6794   & 298  & 100 & 80 & [\citenum{bergerFILESpopc34chol}]$^*$ & [\citenum{ferreira13}] \\
                               &   POPC &64 & 64 & 50\% & 10314  & 298  & 100 & 80 & [\citenum{bergerFILESpopc50chol}]$^*$ & [\citenum{ferreira13}] \\
                               &   POPC &50 & 78 & 61\% & 5782   & 298  & 100 & 80 & [\citenum{bergerFILESpopc60chol}]$^*$ & [\citenum{ferreira13}] \\
CHARMM36\cite{klauda10,lim12}   & POPC   & 128& 0& 0\% & 5120  & 303  & 150 & 100 & [\citenum{charmm36files}]$^*$  & SI  \\
                                & POPC   & 512& 0& 0\% & 23943  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 460& 52& 10\% & 23569  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 436& 76& 15\% & 23331  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 100 & 24 & 19\%  &  4960   & 303 & 200 & 100 & [\citenum{charmm36files20perCHOL}]$^*$ & SI \\
                                & POPC   & 410& 102& 20\% & 20972  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 384& 128& 25\% & 22327  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 332& 180& 35\% & 21340  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 256& 256& 50\% & 20334  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                               & POPC   & 80 & 80 &50\%  &  4496    & 303 & 200 & 100 & [\citenum{charmm36files50perCHOL}]$^*$ & SI \\
MacRog\cite{maciejewski14}     & POPC   & 128 & 0 & 0\% & 6400  & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI \\ 
                          & POPC   & 114  & 14 & 11\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
                          & POPC   & 72   & 56 &  44\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
                             & POPC   & 64  & 64 & 50\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
                             & POPC   & 56   & 72 & 56\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
\end{tabular}
\end{sidewaystable}

@@

\subsubsection{MacRog}
The lipid force field parameters were obtained from the developers and they correspond to the published DPPC parameters~\cite{maciejewski14} with the inclusion of the 
double bond parameters~\cite{kulig15}. 
%This inclusion of unsaturated lipid tails will be published in the near future. 
%\todo{You have recent paper where you use POPC/chol simulations with this model: Kulig et al. BBA 1848 (2015) 422-432 http://dx.doi.org/10.1016/j.bbamem.2014.10.032.
%would this be a correct refence here?}
A bilayer with 288 POPC lipids was hydrated with 12600 TIP3P water~\cite{jorgensen83} molecules ($\sim$44/lipid) and simulated for 100~ns with a time step of 2~fs. Data was saved 
every 10~ps and the first 20~ns of the trajectory was discarded from the analysis. 
All bond lengths were constrained with LINCS~\cite{hess97,hess07}. The temperatures of the lipids and the solvent were separately coupled to the Nos\'{e}--Hoover thermostat~\cite{nose84,hoover85} 
with a target temperature of 310~K and a time constant of 0.4~ps. Semi-isotropical pressure coupling to 1~bar was obtained with the Parrinello--Rahman 
barostat~\cite{parrinello81} with a time constant of 1~ps. PME~\cite{darden93,essman95} was employed to calculate the long-range electrostatic interactions. Lennard-Jones interactions were cut off 
at 1~nm and the dispersion correction was applied to both energy and pressure. A neighbour list with a radius of 1~nm was updated every step.

>>>>

Berger-DPPC-98~\cite{marrink98}    & DPPC &  72 & 2864 & 323 & 60 & 30 & [\citenum{bergerDPPCfiles}]$^*$ & SI \\

@@

CHARMM36~\cite{klauda10}  & DPPC &  72 & 2189 & 323 &  130  & - & -   & [\citenum{klauda10}]$^h$  \\

@@

MacRog~\cite{kulig15b}  & POPC & 288 & 12600 & 310 & 100 &  80 & [\citenum{macrogFILES}]$^*$      & SI  \\

@@

MacRog}~\cite{kulig15b} & POPC & 128 &  6400 & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$  & SI  \\
MacRog~\cite{kulig15b}  & POPC & 288 & 14400 & 310 &  90 &  40 & [\citenum{macrogdehydFILES}]$^*$ & SI  \\

@@

\begin{table}[htb]
\centering
\caption{Simulated single component lipid bilayers with varying hydration levels. The simulation file data sets marked with $^*$ include also part of the trajectory.
$^a$ Water/lipid molar ratio
$^b$ The number of lipid molecules
$^c$ The number of water molecules
$^d$ Simulation temperature
$^e$ The total simulation time
$^f$ Time frames used in the analysis
$^g$ Reference link for the downloadable simulation files
$^h$ Reference for the full simulation details
}\label{systemsDEHYD}
\begin{tabular}{c c c c c c c c c c}
%\hline
Force field & lipid & $^a$n (w/l)   & $^b$N$_{\rm l}$   &  $^c$N$_{\rm w}$ & $^d$T (K)  & $^e$t$_{{\rm sim}}$(ns)  & $^f$t$_{{\rm anal}}$ (ns)& $^g$Files  &  $^h$Details\\
\hline
Berger-POPC-07~\cite{ollila07a}          &   POPC & 57  &128 & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & SI \\
                                        &   POPC & 7  &128 & 896   & 298  & 60 & 50 & [\citenum{bergerDEHYDfiles}]$^*$ & SI \\
Berger-DLPC-13~\cite{kanduc13}          &   DLPC & 28  &72 & 2016  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc28}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 24  &72 & 1728  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc24}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 20  &72 & 1440  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc20}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 16  &72 & 1152  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc16}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 12  &72 & 864  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc12}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 8  &72 & 576  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc8}]$^*$ & [\citenum{kanduc13}] \\
                                 &   DLPC & 4  &72 & 288  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc4}]$^*$ & [\citenum{kanduc13}] \\
CHARMM36\cite{klauda10}          & POPC   & 40 & 128 &  5120   & 303 & 150 & 100  & [\citenum{charmm36files}]$^*$   & SI \\
                                 & POPC   & 31  & 72  &  2242 & 303 & 30 & 20 & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
                               & POPC   & 15 & 72 &  1080  & 303 & 59 & 40 & [\citenum{charmm36files15wPERl}]$^*$ & SI \\
                            & POPC   & 7  & 72  &  504  & 303 & 60 & 20 & [\citenum{charmm36files7wPERl}]$^*$ & SI \\
MacRog\cite{kulig15b}     & POPC   & 50 & 288  & 14400 & 310 & 90 & 40 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                               & POPC   & 25 & 288  & 7200 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 20 & 288  & 5760 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 15 & 288  & 4320 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 10 & 288  & 2880 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 5 & 288   & 1440 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
GAFFlipid\cite{dickson12}      & POPC   & 31& 126  & 3948  & 303 & 137 & 32 & [\citenum{GAFFlipidFILES}]$^*$ & SI  \\
                               & POPC   & 7& 126  & 896   & 303 & 130 & 40 & [\citenum{gaffDEHYDfiles}]$^*$ & SI \\
\end{tabular}
\end{table}

\begin{sidewaystable}[]
\centering
\caption{Simulated lipid bilayers containing cholesterol. The simulation file data sets marked with $^*$ include also part of the trajectory.
$^a$ The number of lipid molecules
$^b$ The number of cholesterol molecules
$^c$ Cholesterol concentration (mol\%)
$^d$ The number of water molecules
$^e$ Simulation temperature
$^f$ The total simulation time
$^g$ Time frames used in the analysis
$^h$ Reference link for the downloadable simulation files
$^i$ Reference for the full simulation details
}\label{systemsCHOL}
\begin{tabular}{c c c c c c c c c c c}
%\hline
Force field & lipid   & $^a$N$_{\rm l}$ & $^b$N$_{\rm chol}$ &$^c$C$_{\rm CHOL}$  &  $^d$N$_{\rm w}$ & $^e$T (K)  & $^f$t$_{{\rm sim}}$(ns)  & $^g$t$_{{\rm anal}}$ (ns)& $^h$Files  &  $^i$Details\\
\hline
Berger-POPC-07~\cite{ollila07a}&   POPC &128 & 0 &0\% & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & [\citenum{ferreira15}] \\
/H\"oltje-CHOL-13~\cite{holtje01,ferreira13}   &    & &  &   &   &  &  &  &  \\
                               &   POPC &120 & 8 & 6\% &7290   & 298  & 100 & 80 & [\citenum{bergerFILESpopc7chol}]$^*$ & [\citenum{ferreira13}] \\
                               &   POPC &110 & 18& 14\% & 8481  & 298  & 100 & 80 & [\citenum{bergerFILESpopc15chol}]$^*$ & [\citenum{ferreira13}]  \\
                               &   POPC &84 & 44 & 34\%  & 6794   & 298  & 100 & 80 & [\citenum{bergerFILESpopc34chol}]$^*$ & [\citenum{ferreira13}] \\
                               &   POPC &64 & 64 & 50\% & 10314  & 298  & 100 & 80 & [\citenum{bergerFILESpopc50chol}]$^*$ & [\citenum{ferreira13}] \\
                               &   POPC &50 & 78 & 61\% & 5782   & 298  & 100 & 80 & [\citenum{bergerFILESpopc60chol}]$^*$ & [\citenum{ferreira13}] \\
CHARMM36\cite{klauda10,lim12}   & POPC   & 128& 0& 0\% & 5120  & 303  & 150 & 100 & [\citenum{charmm36files}]$^*$  & SI  \\
                                & POPC   & 512& 0& 0\% & 23943  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 460& 52& 10\% & 23569  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 436& 76& 15\% & 23331  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 100 & 24 & 19\%  &  4960   & 303 & 200 & 100 & [\citenum{charmm36files20perCHOL}]$^*$ & SI \\
                                & POPC   & 410& 102& 20\% & 20972  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 384& 128& 25\% & 22327  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 332& 180& 35\% & 21340  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 256& 256& 50\% & 20334  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                               & POPC   & 80 & 80 &50\%  &  4496    & 303 & 200 & 100 & [\citenum{charmm36files50perCHOL}]$^*$ & SI \\
MacRog\cite{kulig15b}     & POPC   & 128 & 0 & 0\% & 6400  & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI \\ 
                          & POPC   & 114  & 14 & 11\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
                          & POPC   & 72   & 56 &  44\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
                             & POPC   & 64  & 64 & 50\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
                             & POPC   & 56   & 72 & 56\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
\end{tabular}
\end{sidewaystable}

@@

\subsubsection{MacRog}
The lipid force field parameters were obtained from the developers and the parameters are presented in Refs.~\citenum{maciejewski14,kulig15b}. 
%correspond to the published DPPC parameters~\cite{maciejewski14} with the inclusion of the double bond parameters~\cite{kulig15}. 
%This inclusion of unsaturated lipid tails will be published in the near future. 
%\todo{You have recent paper where you use POPC/chol simulations with this model: Kulig et al. BBA 1848 (2015) 422-432 http://dx.doi.org/10.1016/j.bbamem.2014.10.032.
%would this be a correct refence here?}
A bilayer with 288 POPC lipids was hydrated with 12600 TIP3P water~\cite{jorgensen83} molecules ($\sim$44/lipid) and simulated for 100~ns with a time step of 2~fs. Data was saved 
every 10~ps and the first 20~ns of the trajectory was discarded from the analysis. 
All bond lengths were constrained with LINCS~\cite{hess97,hess07}. The temperatures of the lipids and the solvent were separately coupled to the Nos\'{e}--Hoover thermostat~\cite{nose84,hoover85} 
with a target temperature of 310~K and a time constant of 0.4~ps. Semi-isotropical pressure coupling to 1~bar was obtained with the Parrinello--Rahman 
barostat~\cite{parrinello81} with a time constant of 1~ps. PME~\cite{darden93,essman95} was employed to calculate the long-range electrostatic interactions. Lennard-Jones interactions were cut off 
at 1~nm and the dispersion correction was applied to both energy and pressure. A neighbour list with a radius of 1~nm was updated every step.
